\documentclass{article}
\usepackage[dvips]{graphicx}
\usepackage{a4wide}
\usepackage{amsmath}
\usepackage{euscript}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsopn}

\theoremstyle{definition}
\newtheorem*{definition}{Definition}
\newtheorem{theorem}{Theorem}

\newcommand{\vv}{\ensuremath{\vec{v}}}
\newcommand{\vu}{\ensuremath{\vec{u}}}
\newcommand{\vw}{\ensuremath{\vec{w}}}
\newcommand{\vx}{\ensuremath{\vec{x}}}
\newcommand{\vy}{\ensuremath{\vec{y}}}
\newcommand{\vb}{\ensuremath{\vec{b}}}
\newcommand{\vo}{\ensuremath{\vec{0}}}
\newcommand{\va}{\ensuremath{\vec{a}}}
\newcommand{\ve}{\ensuremath{\vec{e}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%


\begin{document}
\begin{center}
\large{Hints \& Solutions to SUGGESTED PROBLEMS from Sections 1.2 \& 1.3}

\vspace{0.2cm}

Math 251W: Foundations of Advanced Mathematics

\vspace{0.25cm}
\hrule
\end{center}
\begin{center}

\vspace{0.75cm}
\end{center}

\noindent {\it Note: The following are brief solutions or proofs for
selected problems.  Remember, the answer is the least important
part. It's understanding how to get the answer and how to explain
your process which is important.} \vspace{0.5cm}

\noindent 1.2.1:

\noindent (1) Statement.  It is either a nice day or it is not a
nice day

\noindent (2) Not a statement.  Command

\noindent (3) Not a statement.  Question

\noindent (4) Statement.  It is either true or false.

\noindent (5) Statement.

\noindent (6) Statement.

\noindent (7) Not a statement.  "Call me on Thursday" is a command

\vspace{0.5cm}


\noindent 1.2.4:

\noindent (1) If I am eating spaghetti, then I am happy.

\noindent (2) I am happy if and only if I am watching a movie.

\noindent (3) If I am watching a movie or I am eating spaghetti,
then I am happy.

\noindent (4) I am watching a movie, or if I am eating spaghetti
then I am happy.

\noindent (5) If I am watching a movie then I am not happy, and if I
am eating spaghetti then I am not happy.

\noindent (6) I am happy and not watching a movie, if and only if I
am watching a movie or eating spaghetti.

\vspace{0.5cm}

\noindent 1.2.7:

\noindent (4) True.  $\neg D$ is False, $\neg C$ is True, thus $\neg
D \vee \neg C$ is True.

\vspace{0.25cm}

\noindent (5) True.  $D$ is true, and $A$ is true, thus $D\wedge A$
is true.  $B$ is false and $C$ is false, thus $B\wedge C$ is false.
Since $D\wedge A$ is true, it follows that $(D\wedge A)\vee (B\wedge
C)$ is True.

\vspace{0.25cm}

\noindent (6) True.  Since $B$ is false, $A\wedge B$ is false.
However, since $D$ is true, $D\vee(A\wedge B)$ is true, as is
$C\vee(D\vee(A\wedge B))$.

\vspace{0.5cm}


\noindent 1.2.8:

\noindent  (4) True.  Since $X$ is false, $X\rightarrow \neg W$ is
true.  Thus, since both $W$ and $X\rightarrow \neg W$ are true,
$W\rightarrow (X\rightarrow \neg W)$ is true.

\vspace{0.25cm}

\noindent (5) True.  Since $Y$ and $W$ are both true, $Y\rightarrow
W$ is true, and so is $(Y\rightarrow W)\leftrightarrow W$.  Since
$X$ is false, $\neg X$ is true.  Thus, $((Y\rightarrow
W)\leftrightarrow W)\wedge \neg X$ is true.

\vspace{0.25cm}

\noindent  (6) True.  Since $X$ is false and $W$ is true,
$W\rightarrow X$ is False.  Since $Y$ is true, $Z\vee Y$ is true,
and thus $\neg(Z\vee Y)$ is false.  Since the biconditional is true
when both parts are false, the compound statement $(W\rightarrow
X)\leftrightarrow \neg(Z\vee Y)$ is true.

\vspace{0.5cm}

\noindent 1.2.9:

\noindent Let $F$ be the statement "Flora likes fruit," $C$ be the
statement "Flora likes carrots," $N$ be the statement "Flora likes
nuts," and $R$ be the statement "Flora likes rutabagas".  From the
supposition, $F$ and $N$ are true while $C$ and $R$ are false.

\vspace{0.25cm}

\noindent (1) False.  Since $C$ is False, $F\wedge C$ is false.

\vspace{0.25cm}

\noindent (2) True.  Since $N$ is true, $N\vee R$ is true, and since
$C$ is false, $\neg C$ is true.  Thus $(N\vee R)\wedge \neg C$ is
true.

\vspace{0.25cm}

\noindent (3) True.  Since both $F$ and $N$ are true, $F\wedge N$ is
true.  Thu, $C\vee(F\wedge N)$ is true.

\vspace{0.25cm}

\noindent (4) False.  Since both $C$ and $R$ are false, $C\vee R$ is
false.  Thus, $(F\vee N)\wedge (C\vee R)$ is false.

\vspace{0.25cm}

\noindent (5) False.  Since both $C$ and $R$ are false, $C\vee R$ is
false, and thus so to is $F\wedge(C\vee R))$. Thus,
$R\vee(F\wedge(C\vee R))$ is false.

\vspace{0.5cm}

\noindent 1.2.13:

\noindent You can verify each of these with a truth table.  I've
shown this work for part (5), but not for the other parts.

\vspace{0.25cm}

\noindent (1) Neither.

\vspace{0.25cm}

\noindent (2) Tautology

\vspace{0.25cm}

\noindent (3) Contradiction.

\vspace{0.25cm}

\noindent (4) Neither

\vspace{0.25cm}

\noindent (5) Tautology

$\begin{array}{c|c|c|ccccccccccc}L&M&N&(L&\rightarrow&(M&\rightarrow&N))&\rightarrow&(M&\rightarrow&(L&\rightarrow&N))\\\hline
T & T & T & T & T & T & T & T & T & T & T & T & T & T\\
T & T & F & T & F & T & F & F & T & T & F & T & F & F\\
T & F & T & T & T & F & T & T & T & F & T & T & T & T\\
T & F & F & T & T & F & T & F & T & F & T & T & F & F\\
F & T & T & F & T & T & T & T & T & T & T & F & T & T\\
F & T & F & F & T & T & F & F & T & T & T & F & T & F\\
F & F & T & F & T & F & T & T & T & F & T & F & T & T\\
F & F & F & F & T & F & T & F & T & F & T & F & T & F\end{array}$

\vspace{0.25cm}

\noindent (6) Neither

\vspace{0.25cm}

\noindent (7) Contradiction

\vspace{0.5cm}



\noindent 1.2.14:

\noindent (1) $B\rightarrow (B\vee \neg B)$ Tautology

\noindent (2) $(B\vee\neg B)\rightarrow B$ Neither

\noindent (3) $(W\wedge \neg W)\rightarrow E$ Tautology

\noindent (4) $(M\rightarrow C)\wedge (\neg C\wedge M)$
Contradiction

\noindent (5) $(C\vee P)\wedge(P\rightarrow C)$ Neither

\noindent (6) $(G\vee\neg G)\leftrightarrow (B\wedge \neg B)$
Contradiction

\vspace{0.5cm}

\noindent 1.3.1:

\noindent (2)

$\begin{array}{c|c|ccccccccc}P&Q&(P&\rightarrow&Q)&\wedge&(P&\rightarrow&\neg
Q)&\rightarrow&\neg P\\\hline
T & T & T & T & T & F & T & F & F & T & F\\
T & F & T & F & F & F & T & T & T & T & F\\
F & T & F & T & T & T & F & T & F & T & T\\
F & F & F & T & F & T & F & T & T & T & T\\
  &   & 1 & 6 & 2 & 8 & 3 & 7 & 4 & 9 & 5
\end{array}$

\vspace{0.25cm}

\newpage

\noindent (3)

$\begin{array}{c|c|c|ccccccccccc}P&Q&R&(P&\rightarrow&Q)&\rightarrow&(P&\wedge&R)&\rightarrow&(Q&\wedge&R)\\\hline
T & T & T & T & T & T & T & T & T & T & T & T & T & T\\
T & T & F & T & T & T & T & T & F & F & T & T & F & F\\
T & F & T & T & F & F & T & T & T & T & F & F & F & T\\
T & F & F & T & F & F & T & T & F & F & T & F & F & F\\
F & T & T & F & T & T & T & F & F & T & T & T & T & T\\
F & T & F & F & T & T & T & F & F & F & T & T & F & F\\
F & F & T & F & T & F & T & F & F & T & T & F & F & T\\
F & F & F & F & T & F & T & F & F & F & T & F & F & F\\
  &   &   & 1 & 7 & 2 & 11& 3 & 8 & 4 & 10 & 5 & 9 & 6
\end{array}$

\vspace{0.5cm}


\noindent 1.3.2:

\noindent (5)

$\begin{array}{c|c|c|ccccccccccc}P&A&B&P&\rightarrow&(A&\vee&B)&\leftrightarrow&(P&\wedge&\neg
A)&\rightarrow&B\\\hline
T & T & T & T &  T & T & T & T & T & T & F & F &  T & T\\
T & T & F & T &  T & T & T & F & T & T & F & F &  T & F\\
T & F & T & T &  T & F & T & T & T & T & T & T &  T & T\\
T & F & F & T &  F & F & F & F & T & T & T & T &  F & F\\
F & T & T & F &  T & T & T & T & T & F & F & F &  T & T\\
F & T & F & F &  T & T & T & F & T & F & F & F &  T & F\\
F & F & T & F &  T & F & T & T & T & F & F & T &  T & T\\
F & F & F & F &  T & F & F & F & T & F & F & T &  T & F\\
  &   &   & 1 &  8 & 2 & 7 & 3 & 11 & 4 & 9 & 5 &  10 & 6
\end{array}$

\vspace{0.25cm}

\noindent (6)

$\begin{array}{c|c|c|ccccccccccccc}A&B&Q&(A&\vee&B)&\rightarrow&Q&\leftrightarrow&(A&\rightarrow&Q)&\wedge&(B&\rightarrow&Q)\\\hline
T & T & T & T & T & T &  T & T & T & T & T & T &  T & T & T & T\\
T & T & F & T & T & T &  F & F & T & T & F & F &  F & T & F & F\\
T & F & T & T & T & F &  T & T & T & T & T & T &  T & F & T & T\\
T & F & F & T & T & F &  F & F & T & T & F & F &  F & F & T & F\\
F & T & T & F & T & T &  T & T & T & F & T & T &  T & T & T & T\\
F & T & F & F & T & T &  F & F & T & F & T & F &  F & T & F & F\\
F & F & T & F & F & F &  T & T & T & F & T & T &  T & F & T & T\\
F & F & F & F & F & F &  T & F & T & F & T & F &  T & F & T & F\\
  &   &   & 1 & 8 & 2 &  9 & 3 & 13 & 4 & 10 & 5 & 12 & 6 & 11 & 7
\end{array}$

\vspace{0.5cm}

\noindent 1.3.4:

\noindent (1) Let $K$ be the statement "you will kiss me," and let
$J$ be the statement "I will dance a jig."  The question is whether
$(K\rightarrow J)\wedge J$ implies $K$, which it does not.  You can
verify it doesn't using a truth table. The failing case is when $K$
is false, and $J$ is true.  Thus $(K\rightarrow J)\wedge J$ is true
while $K$ is false.

\vspace{0.25cm}

\noindent (3) Let $P$ be the statement "cars pollute" and $T$ be the
statement "we are in trouble."  $(P\rightarrow T)\wedge P
\Rightarrow T$ by modus ponens.

\vspace{0.25cm}

\noindent (4) Let $T$ be the statement "our time is short," $E$ be
the statement "the end is near," and $D$ be the statement "doom is
impending."  The question is whether $(T\vee E)\wedge D$ implies
$E$, which it does not.  Take the case where $T$ is true, $E$ is
false, and $D$ is true.  Thus, $(T\vee E)\wedge D$ is true, but $E$
is false.

\vspace{0.5cm}

\newpage

\noindent 1.3.5:

\noindent (2) Let $S$ be the statement "this shirt has stripes," $L$
be the statement "this shirt has short sleeves," and $B$ be the
statement "this shirt has a band collar."  Thus the statements in
question are $S\wedge(L\vee B)$ and $(S\wedge L)\vee B$ which are
not equivalent.  Take the case when $L$ and $B$ are true but $S$ is
false.  Thus $S\wedge(L\vee B)$ is false while $(S\wedge L)\vee B$
is true.

\vspace{0.25cm}

\noindent (5) Let $R$ be the statement "melons are ripe," and $S$ be
the statement "melons are soft to the touch." $\neg(R\leftrightarrow
S)$ is not equivalent to $(R\wedge S)\vee (\neg R \vee  \neg S)$.
Take the case where $R$ and $S$ are true.  Thus
$\neg(R\leftrightarrow S)$ is false, while $(R\wedge S)\vee (\neg R
\vee  \neg S)$ is true.  The second statement is a tautology, while
the first is not.

\vspace{0.5cm}

\noindent 1.3.9:

\noindent (1) Inverse

\vspace{0.25cm}

\noindent (2) None of these

\vspace{0.25cm}

\noindent (3) Converse

\vspace{0.25cm}

\noindent (4) Contrapositive

\vspace{0.25cm}

\noindent (5) None of these

\vspace{0.25cm}

\noindent (6) Inverse

\vspace{0.5cm}

\noindent 1.3.11:

\noindent (1) It is not Monday or it is not snowing.

\vspace{0.25cm}

\noindent (2) This book is not red and it was not written in 1997.

\vspace{0.25cm}

\noindent (3) Susan does not like to eat figs, or she does not like
to drink prune juice.

\vspace{0.25cm}

\noindent (4) I tell you a joke and you don't smile.

\vspace{0.25cm}

\noindent (5) The play ends on time and the actors are not in good
spirits, or the play does not end on time and the actors are in good
spirits.

\vspace{0.25cm}

\noindent (6) You buy the paint and the room does not get painted.

\end{document}
